spsurveyMaine: select 30 Kennebec county lakes > 150 acres with an unstratified equal probability design
========================================================
# Initial Setup

load the packages needed for the analysis.

```{r setup, include=FALSE, echo=TRUE, cache=FALSE}
 #install (if needed) and load R packages
    libs<-c("spsurvey","sp","maptools","rgdal") #list of packages to load
    installLoad<-function(pck){   #user defined function
      if(!pck%in%installed.packages()){install.packages(pck,repos="http://rweb.quant.ku.edu/cran/")}
        require(pck, character.only = TRUE)
    }
  lapply(libs,function(x) installLoad(x))  #Load/Install require packages
```

# Introduction

Sample design document for Kennebec county lakes > 150 acres.  Linda Bacon provided a shapefile with the lakes and asked for an unstratified equal probability design that includes 30 lakes.  We follow the 1st example in [Finite_Design.pdf](http://cran.r-project.org/web/packages/spsurvey/vignettes/Finite_Design.pdf) by Thomas Kincaid to use the R package spsurvey to conduct a GRTS survey designs for a finite resource. All data and code for this project are available on [github](https://github.com/willbmisled/spsurveyME).

# Data

The sample frame was provided by Linda Bacon as a shapefile with 39 lakes greater than 150 acres in Kennebec County.  The shapefile is in NAD_1983_UTM_Zone_19N.  The first step is reproject to albers equal area. Then the spatialPolygonsDataFrame must be converted to a spatialPointsDataFrame.

```{r projection,echo=TRUE,include=FALSE,cache=TRUE}
  #Input CRS: ESRI NAD_1983_UTM_Zone_19N is EPSG:26919
    NAD83_19N<-CRS('+proj=utm +zone=19 +datum=NAD83 +units=m +no_defs')
  #Output CRS: ESRI North_America_Albers_Equal_Area_Conic
    Albers<-CRS('+proj=aea +x_0=0 +y_0=0 +lon_0=-96 +lat_1=20 +lat_2=60 +lat_0=40 +units=m +datum=NAD83')
  #Read Original Shapefile
    lakesUTM<-readShapePoly('150AcKennebecLakes.shp',proj4string=NAD83_19N)
  #Reproject to Albers
    lakesAlb<-spTransform(lakesUTM,Albers)
  #Convert from spatialPolygonsDataFrame to a spatialPointsDataFrame based on centroids
    lakesAlbPts<-SpatialPointsDataFrame(coordinates(lakesAlb),slot(lakesAlb,"data"), proj4string=Albers)
```

# spsurvey GRTS unstratified, equal probability design 

The finite resource used in the designs is lakes>150acres in Kennebec county Maine.  We use the R package "spsurvey" to develop unstratified, equal probability design for up to 30 lakes. The set.seed function is called so that, if necessary, the designs can be replicated.


```{r SPsurvey,echo=TRUE,include=TRUE,cache=TRUE}
  # Read the attribute table from the shapefile
    att <- read.dbf("lakesAlb") 
  # Call the set.seed function so that the survey designs can be replicate
    set.seed(108)
  # Create the design list
    Equaldsgn <- list(None=list(panel=c(PanelOne=30), seltype="Equal")) #PanelOne=30 is the sample size
  # Select the sample
    Equalsites <- grts(design=Equaldsgn, #design file specified above
                   src.frame="sp.object",#source of the sample frame-in this case an the sp object "lakesAlb"
                   sp.object=lakesAlbPts,  #input sp object
                   shapefile=TRUE)      #whether to output the results as a shapefile
```

# Results

Below is a list of the selected sites.  The siteIDs are assigned by SPsurvey.  The idea is that you select the sites in order of the siteIDs.  If a site cannot be visited for any reason choose the next site from the list.  There is also a rough figure showing all of the lakes as polygons and the selected lakes as the blue centroids.
```{r printResults,echo=TRUE,include=TRUE,cache=TRUE}

  # Print the initial six lines of the survey design
    Equalsites@data[,c('siteID','GNIS_ID','GNIS_Name','AreaSqKm','ReachCode','acres')]
  #create "map" of results
    lakesSelect<-readShapePoints('sample.shp',proj4string=Albers) #read the shapefile of selected lake (points)
    plot(lakesAlb)
    plot(lakesSelect,add=T,pch=19,col='blue')
```


