---
title: "spsurveyME2015"
author: "Bryan Milstead"
date: "Monday, July 27, 2015"
output: pdf_document
---
<!---
use these command instead of the knit icon if you want the data and work loaded into the R workspace
  library(knitr)
    setwd("2015")
    knit('spsurveyME2015.rmd')
-->

##Introduction

* Linda Bacon has asked for a selection of 30 ME lakes for Cyanobacteria monitoring during the summer of 2015 with an unstratified equal probability design.
* The sampling frame is a shapefile (Lakesover150A.shp) from Linda with a set of 45 lakes > 150 acres in Androscoggin, Knox, Lincoln and Sagadahoc counties (Maine)
* The R package spsurvey will be used to select the lakes following the 1st example in <http://cran.r-project.org/web/packages/spsurvey/vignettes/Finite_Design.pdf> by Thomas Kincaid.
* This same approach was used last year.  See: <https://github.com/willbmisled/2014/spsurveyME2014.md>
* The final document for the 2015 draw is available here: <https://github.com/willbmisled/2015/spsurveyME2015.pdf>
* The complete document with code chunks is here: <https://github.com/willbmisled/2015/spsurveyME2015.rmd>

##Workflow

```{r setup, include=FALSE, echo=FALSE, cache=FALSE}
 #install (if needed) and load R packages
    libs<-c("spsurvey","sp","maptools","rgdal") #list of packages to load
    installLoad<-function(pck){   #user defined function
      if(!pck%in%installed.packages()){install.packages(pck,repos="http://rweb.quant.ku.edu/cran/")}
        require(pck, character.only = TRUE)
    }
  lapply(libs,function(x) installLoad(x))  #Load/Install require packages

```

* load r packages:  `r libs`
* read the sample frame (Lakesover150A.shp)
* reproject data from NAD_1983_UTM_Zone_19N to Albers
* Convert from spatialPolygonsDataFrame to a spatialPointsDataFrame based on centroids

```{r projection,echo=TRUE,include=FALSE,cache=FALSE}
  #Input CRS: ESRI NAD_1983_UTM_Zone_19N is EPSG:26919
    NAD83_19N<-CRS('+proj=utm +zone=19 +datum=NAD83 +units=m +no_defs')
  #Output CRS: ESRI North_America_Albers_Equal_Area_Conic
    Albers<-CRS('+proj=aea +x_0=0 +y_0=0 +lon_0=-96 +lat_1=20 +lat_2=60 +lat_0=40 +units=m +datum=NAD83')
  #Read Original Shapefile
    lakesUTM<-readShapePoly('Lakesover150A.shp',proj4string=NAD83_19N)
  #Reproject to Albers
    lakesAlb<-spTransform(lakesUTM,Albers)
  #Convert from spatialPolygonsDataFrame to a spatialPointsDataFrame based on centroids
    lakesAlbPts<-SpatialPointsDataFrame(coordinates(lakesAlb),slot(lakesAlb,"data"), proj4string=Albers)
```

* draw 30 lakes from sample frame with the spsurvey GRTS unstratified, equal probability design 

  ```{r SPsurvey,echo=FALSE,include=FALSE,cache=FALSE}
  # Read the attribute table from the shapefile
    att <- read.dbf("Lakesover150A.dbf") 
  # Call the set.seed function so that the survey designs can be replicated
    set.seed(108)
  # Create the design list
    Equaldsgn <- list(None=list(panel=c(PanelOne=30), seltype="Equal")) #PanelOne=30 is the sample size
  # Select the sample
    Equalsites <- grts(design=Equaldsgn, #design file specified above
                   src.frame="sp.object",#source of the sample frame-in this case an the sp object "lakesAlb"
                   sp.object=lakesAlbPts,  #input sp object
                   shapefile=TRUE)      #whether to output the results as a shapefile
```

# Results

Below is a list of the selected sites.  The siteIDs are assigned by SPsurvey.  The idea is that you select the sites in order of the siteIDs.  If a site cannot be visited for any reason choose the next site from the list.  There is also a rough figure showing all of the lakes as polygons and the selected lakes as the blue centroids.

```{r printResults,echo=FALSE,include=TRUE,cache=FALSE}
    Equalsites@data[,c("siteID","LAKENAME","MIDAS_NUM" )]
  #create "map" of results
    lakesSelect<-readShapePoints('sample.shp',proj4string=Albers) #read the shapefile of selected lake (points)
    plot(lakesAlb)
    plot(lakesSelect,add=T,pch=19,col='blue')
```

